Auditory Scene Similarity Retrieval and Classification with Relevance based Quantization of Scattering Features

### 2016-12-12 paper/structureScene.tex
Edit results section

The same observations are made for the MFCCs for $k\leq5$. However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently. This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless.

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$.}
\label{fig:ASS_2}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/folds.png}
\caption{Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.  Performance on each fold is depicted using a given symbol.
The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
\end{center}
\label{fig:folds}
\end{figure}

We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to the its 2016 edition.

@@

As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
Not only the average accuracy is improved, but the variance between fold-wise accuracies is also reduced, \eg from $16\%$ to $7\%$ in the case of early integration.
This is in accordance with the previous results on similarity retrieval, as well as with other papers comparing mel-frequency coefficients with scattering coefficients on a variety of tasks, ranging from musical genre recognition \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.

\subsubsection*{Early \vs late integration}
Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second.
Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.

Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in temporal integration, namely majority voting. Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$. This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already gained by the classifiers during the training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}).

@@

Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale.
This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.

\begin{table}
\begin{center}
\caption{Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.}
\begin{tabular}{llll}
             & MFCCs         & scattering & log-scattering  \\
             \hline
early, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
early, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
late, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
late, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
\end{tabular}
\end{center}
\label{table:dcase2013}
\end{table}

\begin{table}
\begin{center}
\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).}
\begin{tabular}{lcc}
Features & MFCCs & log-scattering \\
Classifier  & GMMs    & linear SVMs \\
Temporal integration & late (product rule) & late (majority voting) \\
\midrule
beach & $74.6 \pm 18.9$ & $83.5 \pm \phantom{0}7.1$ \\
bus & $58.2 \pm 18.3$ & $88.8 \pm 11.4$ \\
caf\'{e} / restaurant & $85.1 \pm 10.8$ & $64.5 \pm \phantom{0}8.3$ \\
car & $69.1 \pm 20.8$ & $94.9 \pm \phantom{0}6.1$ \\
city center & $89.6 \pm \phantom{0}9.2$ & $91.7 \pm \phantom{0}9.4$ \\
forest path & $72.4 \pm 23.8$ & $93.8 \pm \phantom{0}7.8$ \\
grocery store & $74.1 \pm 13.7$ & $90.9 \pm \phantom{0}6.9$ \\
home & $78.1 \pm 17.7$ & $56.2 \pm 23.9$ \\
library & $65.1 \pm 23.3$ & $82.0 \pm 13.0$ \\
metro station & $85.2 \pm 15.6$ & $96.0 \pm \phantom{0}2.3$ \\
office & $90.8 \pm 16.0$ & $87.5 \pm 21.7$ \\
park & $25.6 \pm 11.0$ & $75.4 \pm \phantom{0}7.4$ \\
residential area & $75.1 \pm 19.4$ & $44.2 \pm 15.0$ \\
train & $34.1 \pm \phantom{0}7.2$ & $58.3 \pm 10.0$ \\
tram & $85.4 \pm 13.9$ & $83.7 \pm 12.2$ \\
\bottomrule
Average & $70.8 \pm \phantom{0}2.6$ & $79.4 \pm \phantom{0}3.0$
\end{tabular}
\end{center}
\label{table:dcase2016}
\end{table}

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate how general our findings are, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$.
Since the LIBLINEAR package enables faster training than LIBSVM, we have used a linear kernel \cite{Hsu2003}.
Our experiments on the DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model. This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}.
Furthermore, our cross-validation analyses confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.

>>>>

The same observations are made for the MFCCs for $k\leq5$. However for a rank $k$ greater than $5$, all the \emph{ob} approaches perform equivalently. This may be due to the fact that, at some point, the MFCCs fail to dissociate the discriminative events of a scene from the non-informative ones, thus making the cluster selection/weighting step useless. \ja{What is the ``at some point'' referring to? In other words, why do we expect this to happen for larger $k$ but not smaller?}

@@

\begin{figure}[t]
\begin{center}
\includegraphics[width=.9\columnwidth]{bw/unsupervised_test1}
\caption{Acoustic scene similarity retrieval (ASSR) in the DCASE 2013 private dataset: precisions at rank $k$ ($p@k$) obtained for scattering coefficients, with and without logarithmic compression, as a function of the rank $k$. \ja{Why not \emph{ob-a} here? Also, the relationship between the different approaches for the regular scattering coefficients seems to be inverted with respect to the log-scattering and the previous figure. That is, here we have \emph{early} $>$ \emph{ob-w} $\approx$ \emph{ob-c} instead of \emph{ob-w} $>$ \emph{ob-w} $>$ \emph{early}.}}
\label{fig:ASS_2}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{bw/folds.png}
\caption{\label{fig:folds}
Classification accuracies over the five folds of the DCASE 2013 private dataset, for various acoustic descriptors and temporal integration strategies.  Performance on each fold is depicted using a given symbol.
The chosen classifier is a support vector machine (SVM) whose kernel is a Gaussian radial basis function (RBF).
The mean accuracies and fold-wise standard deviations of each experiment are reported in Table \ref{table:dcase2013}.}
\end{center}
\end{figure}

We report experiments on acoustic scene classification (ASC), following the original methodology of the DCASE 2013 challenge, and extend our approach to its 2016 edition.

@@

As shown in Figure \ref{fig:folds}, scattering coefficients outperform MFCCs over all but one fold of the DCASE 2013 private dataset when applying early integration, and over all folds when applying late integration.
Not only is the average accuracy improved, but the variance between fold-wise accuracies is also reduced, such as from $16\%$ to $7\%$ in the case of early integration. \ja{Why does this happen?}
This is in accordance with previous results on similarity retrieval, as well as comparisons between mel-frequency coefficients and scattering coefficients for tasks ranging from musical genre recognition and phone segment classification \cite{Anden2014} to environmental sound classification \cite{Salamon2015}.

\subsubsection*{Early \vs late integration}
Late integration, \ie building a prediction by majority voting over $128$ local windows of duration $T=372\,\mathrm{ms}$, outperforms early integration, \ie classifying the acoustic scene with a global window of duration $T=24\,\mathrm{s}$.
This result suggests that, despite its discriminative power at fine scales, the scattering transform is unable to separately characterize distinct acoustic events beyond the scale of about one second. \ja{This is not suggested since the scattering transform is computed at $T=372\,\mathrm{ms}$ and so cannot be expected to capture larger-scale structures. Rather, this says something about the non-stationarity of the audio signals and the fact that we cannot characterize them by some set of summary statistics.}
Therefore, it is beneficial to regard the scattering transform as a mid-level descriptor of acoustic content, not as a high-level descriptor, and adopt a data-driven strategy for the temporal integration of acoustic events within one scene.

Surprisingly, although we ran a large number of experiments to formulate the relevance-based quantization scheme into a supervised setting for acoustic scene classification, none managed to outperform the baseline strategy in \emph{late} temporal integration, namely majority voting. Among such methods, the best result is achieved by the \emph{ob-c} approach with log-compressed scattering coefficients, reaching an accuracy of $66\% \pm0.7$. \ja{Why don't we display these results? Also how is the variance so much smaller here?} This might be due to the fact that the discriminability gained by the relevance-based quantization approach is already achieved by the classifiers during training and the fact the relevance-based quantization approach is equivalent to an early fusion approach, which in the case of the DCASE 2013 has not been identified as beneficial due to the dataset size (see discussion in Section \ref{sec:eili}). \ja{This sentence is very confusing.}

@@

Logarithmic compression provides a small, albeit consistent boost to acoustic scene classification over all folds in the DCASE 2013 private dataset.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale. \ja{Isn't this why we also standardize the coefficients?}
This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).
Again, this is in accordance with our previous findings on acoustic scene similarity retrieval, which are reported in Figure \ref{fig:ASS_2}.

\begin{table}
\begin{center}
\caption{
\label{table:dcase2013}
Classification accuracies on the DCASE 2013 dataset with various settings of features, temporal integration strategies, and support vector machine kernels.
\ja{The relationship between linear and RBF is inverted for the \emph{early} integration scheme compared to \emph{late}. Is this correct?}
}
\begin{tabular}{llll}
             & MFCCs         & scattering & log-scattering  \\
             \hline
\emph{early}, linear  & $54\pm20$   & $62\pm\phantom{0}6$  & $66\pm11$     \\
\emph{early}, RBF     & $47\pm16$  & $56\pm\phantom{0}7$  & $63\pm\phantom{0}6$   \\
\emph{late}, linear  & $60\pm\phantom{0}9$ & $70\pm\phantom{0}8$  & $75\pm\phantom{0}5$   \\
\emph{late}, RBF     & $68\pm10$ & $73\pm\phantom{0}8$  & $\mathbf{78}\pm\phantom{0}6$   \\
\end{tabular}
\end{center}
\end{table}

\begin{table}
\begin{center}
\caption{Class-wise accuracies on the DCASE 2016 dataset for the baseline algorithm (left) and proposed algorithm (right).}
\begin{tabular}{lccc}
Features & MFCCs & MFCCs & log-scattering \\
Classifier  & GMMs & linear SVMs & linear SVMs \\
Variant of \emph{late} integration & product & majority vote & majority vote \\
\midrule
beach & $74.6 \pm 18.9$ & & $83.5 \pm \phantom{0}7.1$ \\
bus & $58.2 \pm 18.3$ & & $88.8 \pm 11.4$ \\
caf\'{e} / restaurant & $85.1 \pm 10.8$ & & $64.5 \pm \phantom{0}8.3$ \\
car & $69.1 \pm 20.8$ & & $94.9 \pm \phantom{0}6.1$ \\
city center & $89.6 \pm \phantom{0}9.2$ & & $91.7 \pm \phantom{0}9.4$ \\
forest path & $72.4 \pm 23.8$ & & $93.8 \pm \phantom{0}7.8$ \\
grocery store & $74.1 \pm 13.7$ & & $90.9 \pm \phantom{0}6.9$ \\
home & $78.1 \pm 17.7$ & & $56.2 \pm 23.9$ \\
library & $65.1 \pm 23.3$ & & $82.0 \pm 13.0$ \\
metro station & $85.2 \pm 15.6$ & & $96.0 \pm \phantom{0}2.3$ \\
office & $90.8 \pm 16.0$ & & $87.5 \pm 21.7$ \\
park & $25.6 \pm 11.0$ & & $75.4 \pm \phantom{0}7.4$ \\
residential area & $75.1 \pm 19.4$ & & $44.2 \pm 15.0$ \\
train & $34.1 \pm \phantom{0}7.2$ & & $58.3 \pm 10.0$ \\
tram & $85.4 \pm 13.9$ & & $83.7 \pm 12.2$ \\
\bottomrule
Average & $70.8 \pm \phantom{0}2.6$ & & $79.4 \pm \phantom{0}3.0$
\end{tabular}
\end{center}
\label{table:dcase2016}
\end{table}

\subsubsection*{Participation to the DCASE 2016 challenge}
In order to evaluate the generality of our findings, we have taken part in the acoustic scene classification task of the 2016 edition of the DCASE challenge.
This edition relies on the TUT 2016 acoustic scenes dataset, which contains $15$ classes of acoustic scenes, and is larger than the DCASE 2013 private dataset by one order of magnitude \cite{Mesaros2016_EUSIPCO}.
Training support vector machines with a Gaussian RBF kernel is computationally demanding for a dataset of such size, especially under a late integration scheme, wherein the number of training instances exceeds $10^5$, we therefore use a linear kernel instead.
Our experiments on the smaller DCASE 2013 dataset, as reported in table \ref{table:dcase2013}, show that the linear kernel gives comparable, albeit slightly worse, classification results than the Gaussian RBF kernel.

Our results on the public part of the DCASE 2016 dataset are reported in Table \ref{table:dcase2016}, compared with a baseline using MFCCs and a generative classifier, namely Gaussian mixture models with $16$ Gaussians per model. This baseline relies on a late integration scheme, wherein the reduction function $\rho$ is a product of likelihoods.
The average accuracy across classes on the public part is brought from $70.8\%$ to $79.4\%$.\footnote{Challenge results on the private part of the dataset are expected to be published in August 2016.}
Furthermore, our cross-validation analysis confirmed that the value $T=372\,\mathrm{ms}$ for the support of the low-pass filter $\boldsymbol{\phi}(t)$ is optimal when applying late integration by majority voting.
